\vspace{-0.8in}Brief Announcement: A Family of Leaderless Generalized-Consensus Algorithms

### 2016-02-12 podc16/paper2.tex
working on correctness

We consider a set of processes $P$ subject to crash-stop faults and communicating by message-passing in an asynchronous network. Processes in $P$ must solve the Generalized Consensus problem~\cite{Lamport05GeneralizeConsensus}. Each process receives proposals for commands of the form $\opn{Propose}(c)$, and must produce responses of the form $\opn{Decide}(\sigma)$, where $\sigma$ is a c-struct. 
The sequence of calls observed must satisfy the non-triviality, consistency, stability, and liveness properties of generalized consensus.
We use the same notation and refer the reader to~\cite{Lamport05GeneralizeConsensus} for a thourough discussion of c-structs and generalized consensus. Briefly, the c-structs are a set, including $\bot$, with an operator $\bullet$ for appending commands to a c-struct. Intuitively, a c-struct represents a set of sequences that are all equivalent up to the ordering of commutative commands.
We say that two commands \textit{commute} when $\sigma\bullet c_1\bullet c_2 =  \sigma\bullet c_2\bullet c_1 $, for every c-struct $\sigma$.
C-structs are partially ordered: $\sigma_1\sqsubseteq \sigma_2$ iff there exists a sequence of commands $\mathcal{D}$ such that $\sigma_2 = \sigma_1\bullet \mathcal{D}$ and $\bot\sqsubseteq \sigma$ for any
$\sigma$. Two c-structs are compatible when they have a
common upper bound.  The consistency property of generalized consensus requires that any two decided c-structs be compatible.

@@

A dependency-set algorithm exposes the input interface $\opn{announce}_p\left( c \right)$, to announce a command $c$, the output interface $\opn{suggest}_p \left( c,\mathcal{D}\right)$, to suggest a set of dependencies $\mathcal{D}$ for $c$, and an input interface $\opn{commit}\left( c,\mathcal{D} \right)$ (meant to observe the map-agreement sub-algorithm), where $p\in P$ is the process executing the call.
Moreover, a dependency-set algorithm must ensure that:
\begin{itemize}[noitemsep,nolistsep]
    \item[-] Safety: (a), for any suggestion $\opn{suggest}_p \left( c,\mathcal{D}\right)$, every command in $\{c\}\cup \mathcal{D}$ must have been announced before; (b), for any two suggestions $\opn{suggest}_p \left( c_1,\mathcal{D}_1\right)$ and $\opn{suggest}_q \left( c_2,\mathcal{D}_2\right)$, if $c_1$ and $c_2$ do not commute then either $c_1\in \mathcal{D}_2$ or $c_2 \in \mathcal{D}_1$.
%    \item If $c$ was announced by $p$, then eventually a dependency set is suggested for $c$ to $p$.
    \item[-] Liveness: (a) If a suggestion $\opn{suggest}_p\left( c \right)$ was made, then eventually a suggestion $\opn{suggest}\left( c,\mathcal{D} \right)_p$ is made, and (b), if a suggestion $\opn{suggest}_p(c,\mathcal{D})$ has been made and no input $\opn{commit}_q\left( c,\mathcal{D'} \right)$ is observed, then eventually a new suggestion $\opn{suggest}_r\left( c,\mathcal{D''} \right)$ for $r\neq p$ is made.
\end{itemize}

An example of implementation of such an algorithm is depicted by the pseudocode of Figures~\ref{fig:deps-set-algo1} and~\ref{fig:deps-set-algo2}. Due to space constraints, our example focuses on satisfying the safety property only. The main idea is the following: a process $p$ announcing a command $c$ broadcasts $c$ to all the processes, and it suggests the dependency-set $deps_p[c]$ for $c$ as the union of the sets $\mathcal{D}$ that it receives as replies from a quorum of $\mathcal{Q}$ processes; on the other hand, a process that receives an announcement for $c$, computes its reply $\mathcal{D}$ for $c$ as the set of commands $d$ that it received so far, and such that $c$ and $d$ do not commute. The choice of the value of $\mathcal{Q}$ must be such that the intersection of any two quorums is not empty. About correctness, we can easily prove that such an implementation satisfies the safety property by noticing that:  any process $p$ inserts a command $d$ in a dependency-set $\mathcal{D}$ that $p$ sent as a reply for a command $c$, only if $p$ previously received $d$. Therefore, every command suggested in a dependency-set must have been announced before. Further, since a process can suggest a set of dependencies $deps_p[c]$ for a command $c$ after it has received a reply for $c$ from at least a quorum of processes, and thanks to the requirement that we enforced on the intersection of quorums, we can conclude that for any two suggestions $\opn{suggest}_p \left( c_1,deps_p[c_1]\right)$ and $\opn{suggest}_q \left( c_2,deps_q[c_2]\right)$,  if $c_1$ and $c_2$ do not commute then either $c_1\in deps_q[c_2]$ or $c_2 \in deps_p[c_1]$.

@@

\subsection{Consistency and Liveness}
Let us now sketch why our generic leaderless algorithm meets the consistency and liveness requirements of generalized consensus. Consider an execution of the algorithm.
\begin{mylemma}\label{lemma:compat}
If $g_p^1$ is the graph $g_p$ at time $t_1$, $g_q^2$ is the graph $g_q$ at time $t_2$, then $g_p^1$ and $g_q^2$ are compatible.
\end{mylemma}
%\vspace{-5pt}
\begin{proof*}
Follows from the safety property (b) of the map-agreement algorithm and definition of the graph $g_p$.
From (b) it follows that if $c$ is in the domain of both $m_p$ and $m_q$ then $m_p\left[ c \right] = m_q\left[ c \right]$.
\end{proof*}
%\vspace{-5pt}
\begin{mylemma}
    The consistency property always holds.
    \label{lemma:consistency}
\end{mylemma}
%\vspace{-5pt}
\begin{proof*}
    By \cref{lemma:compat} and the safety property (b) of the dependency-set algorithm, $g_p^1$ and $g_q^2$ satisfy the premises of \cref{lemma:graph}. Therefore we get that $g_p^1$ and $g_q^2$ are compatible.
\end{proof*}

>>>>

We consider a set of processes $P$ subject to crash-stop faults and communicating by message-passing in an asynchronous network. Processes in $P$ must solve the Generalized Consensus problem~\cite{Lamport05GeneralizeConsensus}. Each process receives proposals for commands of the form $\opn{Propose}(c)$, and must produce responses of the form $\opn{Decide}(\sigma)$, where $\sigma$ is a c-struct. 
The sequence of calls observed must satisfy the non-triviality, consistency, stability, and liveness properties of generalized consensus.
We use the same notation and refer the reader to~\cite{Lamport05GeneralizeConsensus} for a thourough discussion of c-structs and generalized consensus. Briefly, the c-structs are a set, including $\bot$, with an operator $\bullet$ for appending commands to a c-struct. Intuitively, a c-struct represents a set of sequences that are all equivalent up to the ordering of commutative commands.
We say that two commands \textit{commute} when $\sigma\bullet c_1\bullet c_2 =  \sigma\bullet c_2\bullet c_1 $, for every c-struct $\sigma$.
C-structs are partially ordered: $\sigma_1\sqsubseteq \sigma_2$ iff there exists a sequence of commands $\mathcal{D}$ such that $\sigma_2 = \sigma_1\bullet \mathcal{D}$ and $\bot\sqsubseteq \sigma$ for any
$\sigma$. Two c-structs are compatible when they have a
common upper bound. A c-struct contains a command $c$ when it is of the form $\bot\bullet cs$ where $c$ appears in the sequence of commands $cs$. The consistency property of generalized consensus requires that any two decided c-structs be compatible.

@@

A dependency-set algorithm exposes the input interface $\opn{announce}_p\left( c \right)$, to announce a command $c$, the output interface $\opn{suggest}_p \left( c,\mathcal{D}\right)$, to suggest a set of dependencies $\mathcal{D}$ for $c$, and an input interface $\opn{commit}\left( c,\mathcal{D} \right)$ (meant to observe the map-agreement sub-algorithm), where $p\in P$ is the process executing the call.
Moreover, a dependency-set algorithm must ensure that:
\begin{itemize}[noitemsep,nolistsep]
    \item[-] Safety: (a), for any suggestion $\opn{suggest}_p \left( c,\mathcal{D}\right)$, every command in $\{c\}\cup \mathcal{D}$ must have been announced before; (b), for any two suggestions $\opn{suggest}_p \left( c_1,\mathcal{D}_1\right)$ and $\opn{suggest}_q \left( c_2,\mathcal{D}_2\right)$, if $c_1$ and $c_2$ do not commute then either $c_1\in \mathcal{D}_2$ or $c_2 \in \mathcal{D}_1$.
%    \item If $c$ was announced by $p$, then eventually a dependency set is suggested for $c$ to $p$.
    \item[-] Liveness: (a) If a suggestion $\opn{suggest}_p\left( c \right)$ was made, then eventually a suggestion $\opn{suggest}\left( c,\mathcal{D} \right)_p$ is made, and (b), if a suggestion $\opn{suggest}_p(c,\mathcal{D})$ has been made and no input $\opn{commit}_q\left( c,\mathcal{D'} \right)$ is observed, then eventually a new suggestion $\opn{suggest}_r\left( c,\mathcal{D''} \right)$ for $r\neq p$ is made.
\end{itemize}
\vspace{5cm}

@@

\subsection{Satisfying Consistency and Liveness of Generalized Consensus}
Consider an execution of the algorithm.
\begin{mylemma}\label{lemma:compat}
If $g_p^1$ is the graph $g_p$ at time $t_1$, $g_q^2$ is the graph $g_q$ at time $t_2$, then $g_p^1$ and $g_q^2$ are compatible.
\end{mylemma}
\begin{proof*}
Follows from the safety property (b) of the map-agreement algorithm and definition of the graph $g_p$.
From (b) it follows that if $c$ is in the domain of both $m_p$ and $m_q$ then $m_p\left[ c \right] = m_q\left[ c \right]$.
\end{proof*}
\begin{mythm}
    The consistency property always holds.
    \label{lemma:consistency}
\end{mythm}
\begin{proof*}
    By \cref{lemma:compat} and the safety property (b) of the dependency-set algorithm, $g_p^1$ and $g_q^2$ satisfy the premises of \cref{lemma:graph}. Therefore we get that $g_p^1$ and $g_q^2$ are compatible.
\end{proof*}
\begin{mythm}
    If a command $c$ which is GC-proposed is then repeatedly GC-proposed, then a c-struct containing $c$ is eventually decided.
\end{mythm}
\begin{proof*}
    The liveness properties (a) and (b) of the dependency-set algorithm ensures that some dependency sets are repeatedly proposed for $c$ to the map-agreement algorithm, as long as a commit is not observed. Moreover, the map-agreement algorithm ensures that when dependency sets are repeatedly proposed for $c$, then a dependency set will eventually be committed for $c$. Similarly, any dependency of $c$ is eventually committed and $c$ becomes executable, at which points it is included in the next decided c-struct.
\end{proof*}
